Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-04 testbed.tex
more testbed

We can now change all the files in the disk image.

@@

@@

% Set hostname
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash
mac=$(cat /sys/class/net/eth0/address)
hostname=$(grep $mac nodes.csv | cut -f2 -d' ')

@@

@@

@@

>>>>

We can now change all the files in the disk image as desired.

@@

The \ac{Raspi}s should be setup to be as similar as possible. One of those
things that is prefered to be different on the devices is their hostname.
This helps distinguish them from each other. One solution to this in our
approach would be to uniquely change the hostname in each memory card.
This is however not desired because that would either require modifying
the the customized Raspbian lite image before writing the content to the
memory card or to manually change it in the \ac{Raspi}s after the image has
been written to a meomory card.
Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
in our testbed. Those had to be used to put them in a domain name anyway.

Below is a part of the file where we keep track of the MAC addresses and assign
hostnames. This file will be available online and copied to the Raspbian lite
image in a later step under the listed filename. For different \ac{Raspi}s
the MAC addresses has to be changed. The MAC address of a network card can be found
using the command "ifconfig" or "ip addr".

@@

Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
the testbed, it still needs to know which name to give itself.
We do this using a small Bash script:

% Set hostname
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

script_path="$(dirname $(realpath $0))"
mac=$(cat /sys/class/net/eth0/address)
hostname=$(grep $mac ${script_path}/nodes.csv | cut -f2 -d' ')

@@

Line 1 tells the system to interpret the script using Bash. Line 3 gets the
path of the script itself. Line 4 gets the MAC address. This script will be
located in a \ac{Raspi} and thus return the MAC address of itself when called.
Line 4 searches line by line for this mac address.
Line 8-11 assigns the hostname found in "nodes.csv". If the MAC did not exists,
then it will keep the default hostname of the system.

@@

See the following link for information on this procedure:
\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}

@@

See this link:
\url{https://github.com/chesty/overlayroot}
